---
title: "James Stein Quiz"
author: "Naveen Narayanan Meyyappan"
date: "04/12/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question

Compare the risk of the James stein estimator vs the MLE for k =1 to 100 for random thetai, plot it, use 1000 samples at each k to estimate risk., use N(thetai, 1) for the samples. (rnorm(k))

## Answer

```{include=FALSE}
library(ggplot2)
set.seed(1)
```

The James–Stein estimator is a biased estimator of the mean of Gaussian random vectors.
Let us define a function to compute both the James Stein Estimator and MLE estimator. Then we will use 1000 samples to determine the risk at each k(k ranges from 1 to 100). James-Stein estimate is just the sample mean shrunk towards zero.

```{r}
signcheck<-function(x){ # this function is used to check the sign of the input. Returns zero if the input is negative
        if(x>0){
                return(x)
        }
        else{
                return(0)
        }
}

computejsrisk<-function(X,mu,n,k){
        theta_mle=colMeans(X) # theta for mle is the mean of random normal multivariate distribution
        theta_mle_norm = sqrt(sum(theta_mle*theta_mle)) # We will normalise the theta_mle vector
        theta_js = (signcheck(1-((k-2)/(n*(theta_mle_norm*theta_mle_norm))))*theta_mle) # Computing the James Stein Estimator
        risk_js = sum((mu - theta_js)*(mu - theta_js)) # Calculating the risk for the James Stein Estimator
        return(risk_js)
}
computemlerisk<-function(X,mu){
        theta_mle=colMeans(X) # theta for mle is the mean of random normal multivariate distribution
        risk_mle = sum((mu-theta_mle)^2) # Calculating the mle_risk
        return(risk_mle)
}
```

The above two functions are used to compute both the James Stein Risk and MLE Risk.

Now let us simulate the above two functions for n=1000 and k=1:100.

```{r}
n = 1000 # For 1000 samples
k=100 # k varies from 1 to 100
js_risk = {}
mle_risk = {}
for (i in 1:100){
        mu=rep(0,i) # Creating a vector that denotes the ean of the normal distribution which is now set as zero
        val=rnorm((i*n),0,1) # Creating multivariate normal distribution of size n x k
        X=matrix(val,nrow=n,ncol=i) # Converting the vector to a matrix
        mle_val=computemlerisk(X,mu) # Compute mle_risk
        js_val = computejsrisk(X,mu,n,i) # Compute js_risk
        mle_risk=c(mle_risk,mle_val) # Append the results for each k
        js_risk=c(js_risk,js_val)
}
```

The vector mle_risk and js_risk are used to store the result. Now we will plot the result to compare the James Stein Estimator and MLE Estimator.

```{r}
library(ggplot2)
xaxisvalues=1:100
result=data.frame(xaxisvalues,mle_risk,js_risk)
ggplot() + geom_line(data = result, aes(x = result$xaxisvalues, y = result$mle_risk, colour = "MLE Risk")) + geom_line(data = result, aes(x = result$xaxisvalues, y = result$js_risk, color = "James Stein Risk")) + labs(title="JS Estimator Risk and MLE Estimator Risk", x="K -->", y="Risk -->") +  theme(plot.title = element_text(hjust = 0.5)) + scale_colour_manual("", breaks = c("MLE Risk", "James Stein Risk"),values = c("red", "green", "blue"))
```

## References
1. https://en.wikipedia.org/wiki/James%E2%80%93Stein_estimator
2. https://www.naftaliharris.com/blog/steinviz/















