---
title: "Smooth Regression Assignment"
author: "Naveen Narayanan Meyyappan"
date: "05/12/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question

Construct code for bootstrapping the Press statistic, bootstrapping along the rows. 

And add the press statistic into each the gaussian smooths

Construct confidence intervals for each press statistic for the gaussian smooths.

## Answer

Smooth regression involves drawing a smooth curve on a scatter diagram to summarize a relationship, in a fashion that makes few assumptions initially about the form or strength of the relationship.  It is related to (and is a special case of) nonparametric regression, in which the objective  is to represent the relationship between a response variable and one or more predictor variables, again in way that makes few assumptions about the form of the relationship.  In other words, in contrast to "standard" linear regression analysis, no assumption is made that the relationship is represented by a straight line (although one could certainly think of a straight line as a special case of nonparametric regression). 

Another way of looking at scatter diagram smoothing is as a way of depicting the "local" relationship between a response variable and a predictor variable over parts of their ranges, which may differ from a "global" relationship determined using the whole data set.  (And again, the idea of "local" as opposed to "global" relationships has an obvious geographical analogy.)

Let us define the functions for each of the smooth curves.
```{r}
NOAA1=read.csv("NOAA+GISS.csv")
my.hat.w<-function(x,wt){
        x1<-cbind(1,x)
        x1%*%solve(t(x1)%*%diag(wt)%*%x1)%*%t(x1)%*%(diag(wt))
}
bin.mean<-function(x,y,nbin,xcol=2)
{
        o1<-order(x)
        x1<-x[o1]
        y1<-y[o1]
        r1<-range(x)
        inc<-(r1[2]-r1[1])/nbin
        yvec<-NULL
        smat<-NULL
        for(i in 1:nbin){
                bin.low<-r1[1]+(i-1)*inc
                bin.high<-r1[1]+i*inc
                
                I1<-x1>=bin.low
                if(i<nbin){
                        I2<-x1<bin.high
                }else{
                        I2<-x1<=(bin.high+200)
                }
                I3<-as.logical(I1*I2)
                yval<-mean(y1[I3])
                n1<-sum(I3)
                matdum<-NULL
                for(i in 1:n1){
                        matdum<-rbind(matdum,I3*1/n1)
                }
                smat<-rbind(smat,matdum)
                yvec<-c(yvec,rep(yval,n1))
        }
        n99<-length(x1)
        dferror<-length(x1)-sum(diag(2*smat-smat%*%(t(smat))))
        delta1<-sum(diag(t(diag(n99)-smat)%*%(diag(n99)-smat)))
        R<-t(diag(n99)-smat)%*%(diag(n99)-smat)
        delta2<-2*sum(diag(R%*%R))
        lines(x1,yvec,col=xcol)
        ypred<-y1
        ypred<-smat%*%y1
        resid<-y-ypred
        pressvec<-sum((resid/(1-diag(smat)))^2) 
        list(smat=smat,df=sum(diag(smat)),dferror=dferror,delta1=delta1,delta2=delta2,resid=resid,pred=ypred,press=pressvec)
        
}
gauss.mean<-function(x,y,lambda,xcol=3,do.plot=T)
{
        o1<-order(x)
        x1<-x[o1]
        y1<-y[o1]
        r1<-range(x)
        smat<-NULL
        n1<-length(x1)
        for(i in 1:n1){
                v1<-dnorm(x1,x1[i],lambda)
                v1<-v1/sum(v1)
                smat<-rbind(smat,v1)
        }
        yhat<-smat%*%y1
        if(do.plot){
                lines(x1,yhat,col=xcol)
        }
        n99<-length(x1)
        dferror<-length(x1)-sum(diag(2*smat-smat%*%(t(smat))))
        delta1<-sum(diag(t(diag(n99)-smat)%*%(diag(n99)-smat)))
        R<-t(diag(n99)-smat)%*%(diag(n99)-smat)
        delta2<-2*sum(diag(R%*%R))
        resid<-y1-smat%*%y1
        ypred<-y1
        ypred[o1]<-smat%*%y1
        pressvec<-sum((resid/(1-diag(smat)))^2) 
        list(smat=smat,df=sum(diag(smat)),dferror=dferror,delta1=delta1,delta2=delta2,resid=resid,pred=ypred,press=pressvec)
}
gauss.mean.trunc<-function(x,y,lambda,nnn,xcol=5,do.plot=T)
{
        o1<-order(x)
        x1<-x[o1]
        y1<-y[o1]
        r1<-range(x)
        smat<-NULL
        n1<-length(x1)
        trunc.val<-n1-nnn
        for(i in 1:n1){
                v1<-dnorm(x1,x1[i],lambda)
                o2<-order(v1)
                thresh<-v1[o2[trunc.val]]
                v1<-v1*(v1>thresh)
                v1<-v1/sum(v1)
                smat<-rbind(smat,v1)
        }
        yhat<-smat%*%y1
        if(do.plot){
                lines(x1,yhat,col=xcol)
        }
        n99<-length(x1)
        dferror<-length(x1)-sum(diag(2*smat-smat%*%(t(smat))))
        delta1<-sum(diag(t(diag(n99)-smat)%*%(diag(n99)-smat)))
        R<-t(diag(n99)-smat)%*%(diag(n99)-smat)
        delta2<-2*sum(diag(R%*%R))
        resid<-y1-smat%*%y1
        ypred<-y1
        ypred[o1]<-smat%*%y1
        pressvec<-sum((resid/(1-diag(smat)))^2) 
        list(smat=smat,df=sum(diag(smat)),dferror=dferror,delta1=delta1,delta2=delta2,resid=resid,pred=ypred,press=pressvec)
        
}
gauss.reg<-function(x,y,lambda,xcol=4,do.plot=T)
{
        o1<-order(x)
        x1<-x[o1]
        y1<-y[o1]
        r1<-range(x)
        smat<-NULL
        n1<-length(x1)
        for(i in 1:n1){
                v1<-dnorm(x1,x1[i],lambda)
                v1<-v1/sum(v1)
                H1<-my.hat.w(x1,v1)
                smat<-rbind(smat,H1[i,])
        }
        yhat<-smat%*%y1
        if(do.plot){
                lines(x1,yhat,col=xcol)
        }
        n99<-length(x1)
        dferror<-length(x1)-sum(diag(2*smat-smat%*%(t(smat))))
        delta1<-sum(diag(t(diag(n99)-smat)%*%(diag(n99)-smat)))
        R<-t(diag(n99)-smat)%*%(diag(n99)-smat)
        delta2<-2*sum(diag(R%*%R))
        resid<-y1-smat%*%y1
        ypred<-y1
        ypred[o1]<-smat%*%y1
        pressvec<-sum((resid/(1-diag(smat)))^2) 
        list(smat=smat,df=sum(diag(smat)),dferror=dferror,delta1=delta1,delta2=delta2,resid=resid,pred=ypred,press=pressvec)
}
gauss.reg.trunc<-function(x,y,lambda,nnn,xcol=6,do.plot=T)
{
        o1<-order(x)
        x1<-x[o1]
        y1<-y[o1]
        r1<-range(x)
        smat<-NULL
        n1<-length(x1)
        trunc.val<-n1-nnn
        for(i in 1:n1){
                v1<-dnorm(x1,x1[i],lambda)
                o1<-order(v1)
                thresh<-v1[o1[trunc.val]]
                v1<-v1*(v1>thresh)
                v1<-v1/sum(v1)
                H1<-my.hat.w(x1,v1)
                smat<-rbind(smat,H1[i,])
        }
        yhat<-smat%*%y1
        if(do.plot){
                lines(x1,yhat,col=xcol)
        }
        n99<-length(x1)
        dferror<-length(x1)-sum(diag(2*smat-smat%*%(t(smat))))
        delta1<-sum(diag(t(diag(n99)-smat)%*%(diag(n99)-smat)))
        R<-t(diag(n99)-smat)%*%(diag(n99)-smat)
        delta2<-2*sum(diag(R%*%R))
        resid<-y1-smat%*%y1
        ypred<-y1
        ypred[o1]<-smat%*%y1
        pressvec<-sum((resid/(1-diag(smat)))^2) 
        list(smat=smat,df=sum(diag(smat)),dferror=dferror,delta1=delta1,delta2=delta2,resid=resid,pred=ypred, press=pressvec)
}
```

Let us plot the different smooth regression curves.

```{r}
plot(NOAA1[,3],NOAA1[,2],xlab="temperature rise",ylab="rate of billion dollar weather disasters")
dum<-bin.mean(NOAA1[,3],NOAA1[,2],6)
dum<-gauss.mean(NOAA1[,3],NOAA1[,2],.063)$df
gauss.reg(NOAA1[,3],NOAA1[,2],.078,do.plot=T)$df
gauss.mean.trunc(NOAA1[,3],NOAA1[,2],.063,20,do.plot=T)$df
gauss.reg.trunc(NOAA1[,3],NOAA1[,2],.08,17,do.plot=T)$df
lines(lowess(NOAA1[,3],NOAA1[,2]),col=7)
lines(smooth.spline(NOAA1[,3],NOAA1[,2]),col=8)
```

We have to bootstrap the press vector of different smooth regression curves and also plot the density of the press vector

1. Bootstrapping Gauss.mean function
```{r}
bootstrap<-function(x=NOAA1[,3],y=NOAA1[,2],nboot=1000){
        plot(x,y)
        dum.reg<-gauss.reg(x,y,.2,do.plot=T)
        lstr<-lsfit(x,y)
        abline(lstr)
        pressvec0<-dum.reg$press
        pressvec<-NULL
        for(i in 1:nboot){
                v1<-sample(length(x),length(x),replace=T)
                x1<-x[v1]
                y1<-y[v1]
                dum.reg1<-gauss.reg(x1,y1,.2,do.plot=T)
                pressvec1<-dum.reg1$press
                pressvec<-c(pressvec,pressvec1)
        }
        error <- qt(0.975,df=length(pressvec)-1)*sd(pressvec)/sqrt(length(pressvec))
        LC <- mean(pressvec)-error
        UC <- mean(pressvec)+error
        list(pressvec=pressvec,press=dum.reg$press, CI=list(LC,UC))
}
result=bootstrap()
d<-density(result$pressvec)
plot(d, main="Density of Press Vector for Gauss.mean",xlab="",ylab="")
```

2. Bootstrapping Gauss.mean Function

```{r}
bootstrap<-function(x=NOAA1[,3],y=NOAA1[,2],nboot=1000){
        plot(x,y)
        dum.reg<-gauss.mean(x,y,.2,do.plot=T)
        lstr<-lsfit(x,y)
        abline(lstr)
        pressvec0<-dum.reg$press
        pressvec<-NULL
        for(i in 1:nboot){
                v1<-sample(length(x),length(x),replace=T)
                x1<-x[v1]
                y1<-y[v1]
                dum.reg1<-gauss.mean(x1,y1,.2,do.plot=T)
                pressvec1<-dum.reg1$press
                pressvec<-c(pressvec,pressvec1)
        }
        error <- qt(0.975,df=length(pressvec)-1)*sd(pressvec)/sqrt(length(pressvec))
        LC <- mean(pressvec)-error
        UC <- mean(pressvec)+error
        list(pressvec=pressvec,press=dum.reg$press, CI=list(LC,UC))
}
result=bootstrap()
d<-density(result$pressvec)
plot(d, main="Density of Press Vector for Gauss.mean",xlab="",ylab="")
```

3. Bootstrapping Bin.mean Function

```{r}
bootstrap<-function(x=NOAA1[,3],y=NOAA1[,2],nboot=1000){
        plot(x,y)
        dum.reg<-bin.mean(x,y,.2)
        lstr<-lsfit(x,y)
        abline(lstr)
        pressvec0<-dum.reg$press
        pressvec<-NULL
        for(i in 1:nboot){
                v1<-sample(length(x),length(x),replace=T)
                x1<-x[v1]
                y1<-y[v1]
                dum.reg1<-bin.mean(x1,y1,.2)
                pressvec1<-dum.reg1$press
                pressvec<-c(pressvec,pressvec1)
        }
        error <- qt(0.975,df=length(pressvec)-1)*sd(pressvec)/sqrt(length(pressvec))
        LC <- mean(pressvec)-error
        UC <- mean(pressvec)+error
        list(pressvec=pressvec,press=dum.reg$press, CI=list(LC,UC))
}
result=bootstrap()
d<-density(result$pressvec)
plot(d, main="Density of Press Vector for Bin.mean",xlab="",ylab="")
```


## References 

1. http://geog.uoregon.edu/bartlein/old_courses/geog414f03/lectures/lec05.htm
